PyPy's Approach to Virtual Machine Construction

### 2006-07-12 talk/dls2006/paper.tex
short discussion of RPython features vs Python, and an example of how RPython code is quite like idiomatic Python.

There are two major components in PyPy:
%
\begin{enumerate}
\item the \textit{Standard Interpreter}: an implementation of the Python programming
language, mostly complete and compliant with the current version of the
language, Python 2.4.
\item the \textit{Translation Process}: a translation tool-suite whose goal is to
compile subsets of Python to various environments.
\end{enumerate}
%
In particular, we have defined a subset of the Python language called
``restricted Python'' or RPython.  This sublanguage is not restricted
syntactically, but only in the way it manipulates objects of different
types.  The restrictions are a compromise between the expressivity and
the need to statically infer enough type information to generate efficient code.
The purpose of the translation tool-suite is to compile such
RPython programs to a variety of different platforms.

>>>>

There are two major components in PyPy:
%
\begin{enumerate}
\item the \textit{Standard Interpreter}: an implementation of the Python programming
language, mostly complete and compliant with the current version of the
language, Python 2.4.
\item the \textit{Translation Process}: a translation tool-suite whose goal is to
compile subsets of Python to various environments.
\end{enumerate}
%
In particular, we have defined a subset of the Python language called
``restricted Python'' or RPython.  This sublanguage is not restricted
syntactically, but only in the way it manipulates objects of different
types.  The restrictions are a compromise between the expressivity and
the need to statically infer enough type information to generate
efficient code.  RPython still supports inheritance but limited to
single inheritance with some mix-in support, dynamic dispatch, to some
extent keywords arguments and varargs, first-class function and class
values, limited use of bound methods, runtime 'isinstance' and type
queries, but no runtime reflection, further bindings in class and
global namespaces are assumed constant. RPython code can be run on a
Python interpreter without severe semantics mismatches, figure
\ref{fig_create_frame} shows some RPython code from the
\textit{Standard Interpreter}, it is still quite idiomatic Python code
using the built-in dictionary type and first-class class values
instantiation.

\begin{figure}
\begin{verbatim}
  def get_frame_class(self):
    # select the appropriate kind of frame
    if not frame_classes:
      setup_frame_classes()   # lazily
    choose = 0
    if self.co_cellvars or self.co_freevars:
      choose |= NESTED
    if self.co_flags & CO_GENERATOR:
      choose |= GENERATOR
    Frame = frame_classes[choose]
    return Frame

  def create_frame(self, space, 
        w_globals, closure=None):
    return self.get_frame_class()(space, self, 
              w_globals, closure)
\end{verbatim}
\caption{methods from Python bytecode class to instantiate frames.}
\label{fig_create_frame}
\end{figure}

The purpose of the translation tool-suite is to compile such RPython
programs to a variety of different platforms.
